### 2018-09-28 paper.tex
Wip

@@

A \emph{machine learning model} is challenging to define as models
have different contexts. In the most general context, models are
mathematical concepts or language that represents a system
(needref). In the context of machine learning, models may be
represented by computer source code, either explicitly as declared
structures, or implicitly by way of executable instructions that
perform model related tasks. A \emph{trained model} is a model that
has learned some representation of a system through the process of
computer based optimization.

For the purpose of our discussion, a machine learning model is a
software representation of a specific mathetical concept that can be
trained by a computer. Models may be in various states:
\emph{uninitialized}, \emph{initialized}, \emph{partially trained},
\emph{fully trained}. Regardless of state, the distinguishing feature
of machine learning models for our purpose is that they support
runnable computer operations.

@@

A common approach to sharing and reusing machine learning models is a
\emph{model zoo}---an online repository of reusable models that
includes source code, examples, documentation, and in some cases
pretrained models. The \href{Caffe Model Zoo}{Caffe Model Zoo} was one
of the first model zoos and has inspired serveral others (ref).

@@

\subsubsection{Software packages and source coderepositories}

@@

Guild AI also supports the well established practice of package
maintenence, where a third party can create usable packages without
necessarily coordinating with model developers. This is discussed in
detail in Section \ref{sec:package-maintainer}.

As with traditional software packages and source code repositories,
Guild AI supports a federated approach to software distriution. Users
are free to use Guild AI to create and distribute reusable models in a
variety of ways. Refer to Section \ref{sec:use-cases} for coverage of
various use case.

@@

\begin{lstlisting}[
    caption=Simple Guild file example,
    label={lst:guild-file}]
model: mnist
description:
  Logistic regression classifier
  of MNIST digits
operations:
  train:
    main: train_mnist
    flags:
      batchs-size: 32
      train-steps: 1000
  evaluate:
    main: eval_mnist
\end{lstlisting}

@@

The \verb|train| operation---the act of training a model from
scratch---for example, is an operation that most models
support. However, the \verb|transfer-learn| operation---the act of
transfering knowledge learned in one domain to another domain---is a
form of training, but frequently appears in Guild AI models as an
operation distinct from \verb|train|.

@@

@@

A \emph{developer} may encourage others to use her model by
documenting the steps needed to install, adapt, and use it. She may be
motivated to simplify the user experience so that her work is more
easily and succesfully applied. The benefit to her may be quality
feedback from expert users, improved reputation, a sense of
contribution, or her colleagues may simply expect that work be
documented and easy to use.

While documenting the steps needed to use a model is undoubtedly
helpful, the process is still burdensome for both the developer and
her users. Guild AI can be used to simultaenously document model use
but also directly support its use through model operations.

@@

\begin{enumerate}
   \item Install required packages
   \item Download the raw data
   \item Prepare the raw data for training
   \item Train the model
   \item Export the trained model for deployment
\end{enumerate}

While these steps provide invaulable information, they still rely on
the user to carefully execute them and troubleshoot issues as they
arise. The process takes time and is subject to error at any step.

@@

The flip side of the use case outline in
Section~\ref{sec:publish-model} is that of the \emph{user}. A user is
interesting in taking the work of a model developer and applying it,
either by adapating her model to a new application or by integrating
it as into another project.

@@

Complex software ecosystems such as GNU/Linux have formalized the role
of a software \emph{packager} (or \emph{package maintainer}). The goal
of a packager is to improve the value of a software distribution by
creating high quality software packages that users find
useful. Packers are often independent of the upstream software
projects they package, serving as intermediaries betweeen the software
developers and distribution users.

@@

Given a Guild file with such a package definition, Guild AI will
generate a Python wheel distribution that can be used to install any
models defined in the package.

\subsection{Project Models}

@@

@@

@@

@@

\subsection{Exporting Model Artifacts}

@@

Information about one or more runs may be displayed using the
\verb|guild runs info| command. By default, high level run information
such as the run ID, operation, start time, command, exit status, and
operting system pid are displayed. Additional command options may be
used to view more run information including files, flag values,
dependencies, and operation output.

@@

Guild AI makes use of the Click Python library (ref) to implement its
core command line processingn support.

@@

Guild AI packages are valid Python wheel distributions. The underlying
libraries used to install Guild packages are the same as those used by
the \verb|pip| pgrogram to install PyPI packages. As a result, Guild
AI can be used to install PyPI packages and pip can be used to install
Guild AI packages---the process is the same in both cases.

@@

Guild View is impemented using the VueJS library (ref) for the front
end and Werkzeug (ref) for the HTTP API back end.

@@

Guild AI is freely available under the Apach 2.0 open source
license. It has been under active development for the last two years
and undergone significant revision and updates during that time based
on user feedback and real world application development.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Typography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Delete this to end of doc
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Verbatim}

{\footnotesize
\begin{verbatim}
dvips -Ppdf -tletter -G0 -o paper.ps paper.dvi
ps2pdf paper.ps
another line yo
\end{verbatim}}

You can copy the file by running \verb|cp foo bar| where \textit{foo}
is some file you want to copy and \textit{bar} is something else.

\subsection{Italics}

\textit{Proceedings of the $\mathit{2}^{nd}$ SysML Conference}

And \emph{this is for emphasis!}

\subsection{Bold}

I will \textbf{make this bold} yo.

\subsection{Teletype}

Set the value to \texttt{accepted} and everything will be okay.

\subsection{Dots}

ldots: \ldots

dots: \dots

\subsection{Less than, greater than}

\textless{}email@domain.com\textgreater{}

\subsection{Footnotes}

You can use footnotes\footnote{Footnotes should be complete
  sentences.} which will be super awesome. Don't ask me how it works.

\subsection{Tables}

\begin{table}[t]
  \caption{Classification accuracies for naive Bayes and flexible
    Bayes on various data sets.}
  \label{sample-table}
  \begin{center}
    \begin{small}
      \begin{sc}
        \begin{tabular}{lcccr}
          \toprule
          Data set & Naive & Flexible & Better? \\
          \midrule
          Breast    & 95.9$\pm$ 0.2& 96.7$\pm$ 0.2& $\surd$ \\
          Cleveland & 83.3$\pm$ 0.6& 80.0$\pm$ 0.6& $\times$\\
          Glass2    & 61.9$\pm$ 1.4& 83.8$\pm$ 0.7& $\surd$ \\
          Credit    & 74.8$\pm$ 0.5& 78.3$\pm$ 0.6&         \\
          Horse     & 73.3$\pm$ 0.9& 69.7$\pm$ 1.0& $\times$\\
          Meta      & 67.1$\pm$ 0.6& 76.5$\pm$ 0.5& $\surd$ \\
          Pima      & 75.1$\pm$ 0.6& 73.9$\pm$ 0.5&         \\
          Vehicle   & 44.9$\pm$ 0.6& 61.5$\pm$ 0.4& $\surd$ \\
          \bottomrule
        \end{tabular}
      \end{sc}
    \end{small}
  \end{center}
  \vskip -0.1in
\end{table}

Table \ref{software-vs-models} compares traditional software with
reusable models.

\begin{table}[h!]
  \begin{center}
    \caption{Traditional software vs reusable models.}
    \label{software-vs-models}
    \begin{small}
      \begin{tabular}{l|l}
        \toprule
        \textbf{Traditional software} & Installed code \\ \hline
        \textbf{Reusable models} & Installed code + trainable on novel data \\
        \bottomrule
      \end{tabular}
    \end{small}
  \end{center}
\end{table}

\subsection{Citations}

I really like TensorFlow \cite{tensorflow2015-whitepaper}.

>>>>

\usepackage{enumitem}

\setlist[description]{
  itemsep=4pt,
  labelindent=1em,
  leftmargin=1em,
  topsep=4pt,
}

\setlist[enumerate]{
  itemsep=4pt,
  topsep=4pt,
}

\setlist[itemize]{
  itemsep=4pt,
  topsep=4pt,
}

@@

A \emph{machine learning model} is challenging to define as models
have different contexts. In the most general context, models are
mathematical concepts or language that represents a system (ref). In
the context of machine learning, models may be represented by computer
source code, either explicitly as declared structures, or implicitly
by way of executable instructions that perform model related tasks. A
\emph{trained model} is a model that has learned some representation
of a system through the process of computer based optimization.

For the purpose of our discussion, a machine learning model is a
software representation of a specific mathematical concept that can be
trained by a computer. Models may be in various states:
\emph{uninitialized}, \emph{initialized}, \emph{partially trained},
\emph{fully trained}. Regardless of state, the distinguishing feature
of machine learning models for our purpose is that they support
runnable computer operations.

@@

A common approach to sharing and reusing machine learning models is a
\emph{model zoo}---an online repository of reusable models that
includes source code, examples, documentation, and in some cases
pretrained models. The \href{Caffe Model Zoo}{Caffe Model Zoo} was one
of the first model zoos and has inspired several others (ref).

@@

\subsubsection{Software packages and source code repositories}

@@

Guild AI also supports the well established practice of package
maintenance, where a third party can create usable packages without
necessarily coordinating with model developers. This is discussed in
detail in Section \ref{sec:package-maintainer}.

As with traditional software packages and source code repositories,
Guild AI supports a federated approach to software distribution. Users
are free to use Guild AI to create and distribute reusable models in a
variety of ways. Refer to Section \ref{sec:use-cases} for coverage of
various use case.

@@

\begin{lstlisting}[
    caption=Simple Guild file example,
    label={lst:guild-file}]
model: mnist
description:
  Logistic regression classifier
  of MNIST digits
operations:
  train:
    main: train_mnist
    flags:
      batch-size: 32
      train-steps: 1000
  evaluate:
    main: eval_mnist
\end{lstlisting}

@@

Below are some common operation names:

\begin{description}
\item[prepare] prepare a dataset for training
\item[train] train a model from scratch
\item[transfer-learn] train a model using transfer learning
\item[finetune] finetune a trained model
\item[evaluate] evaluate a model
\item[classify] use a trained model to classify inputs
\item[compress] compress a trained model (e.g. using quantization,
  etc.)
\item[export-and-freeze] export a frozen inference
  graph\footnote{TensorFlow specific}
\item[tflite] generate a TF Lite file from a frozen
  model\footnotemark[\value{footnote}]
\end{description}

This list illustrates that model operations can be used to perform a
host of tasks beyond training and evaluation. In fact, well designed
models can encompass all of the operations associated with a machine
learning application development pipeline, from data acquisition and
preparation to application deployment and monitoring.

@@

Below is an example of using a flag to set the \emph{learning rate}
for a \emph{train} operation.

{\footnotesize
\begin{verbatim}
 $ guild run train learning-rate=0.0001
\end{verbatim}}

@@

A \emph{developer} may encourage others to use her model by
documenting the steps needed to install, adapt, and use it. She may be
motivated to simplify the user experience so that her work is more
easily and successfully applied. The benefit to her may be quality
feedback from expert users, improved reputation, a sense of
contribution, or her colleagues may simply expect that work be
documented and easy to use.

While documenting the steps needed to use a model is undoubtedly
helpful, the process is still burdensome for both the developer and
her users. Guild AI can be used to simultaneously document model use
but also directly support its use through model operations.

@@

\setlength{\parindent}{1em}
\begin{enumerate}
\item Install required packages
\item Download the raw data
\item Prepare the raw data for training
\item Train the model
\item Export the trained model for deployment
\end{enumerate}
\setlength{\parindent}{0em}

While these steps provide invaluable information, they still rely on
the user to carefully execute them and troubleshoot issues as they
arise. The process takes time and is subject to error at any step.

@@

The flip side of the use case outline in
Section~\ref{sec:publish-model} is that of the \emph{user}. A user is
interesting in taking the work of a model developer and applying it,
either by adapting her model to a new application or by integrating
it as into another project.

@@

Complex software ecosystems such as GNU/Linux have formalized the role
of a software \emph{packager} (or \emph{package maintainer}). The goal
of a packager is to improve the value of a software distribution by
creating high quality software packages that users find
useful. Packers are often independent of the upstream software
projects they package, serving as intermediaries between the software
developers and distribution users.

@@

Given a Guild file with a package definition, develolpers can use
Guild AI to generate a Python wheel distribution, which in turn can be
used by users to install package models. Developers can also upload
packages to PyPI to make the package available for search and
installation.

Guild AI models are installed like any other Python package. Users may
use Guild AI to install the package:

Alternatively, a user may use pip:

{\footnotesize
\begin{verbatim}
 $ pip install sample-project
\end{verbatim}}

Guild AI package requirements may be included in requirements files
(e.g. \verb|requirements.txt|) or as dependencies of other Python
package distributions.

Package models are convenient for using models and their related
functionality when a user doesn't need to modify the model source
code. To use models and also have the option of modifying their source
code, a user should use project models (Section
\ref{sec:project-models}).

\subsection{Project Models}
\label{sec:project-models}

Project models are defined in a project Guild file---i.e. a file named
\verb|guild.yml| that resides in the top level of a projet
directory. Users may run operations on project models by running them
from within the project directory or by refering the project directory
in the operation specified for the \verb|run| command.

Project models are used by users or developers who expect to modify
project files.

Unlike packaged models, project models do not need to be installed
using a command---they are available from the project itself. A user
need only change to the project directory to run model operations.

@@

A user may list available models using the \verb|guild models|
command. This command lists packaged models and, if the command is run
within a project, project models.

Each model has one or more operations, which can be listed using the
\verb|guild ops| command.

The ability to discover available models and operations is a key
feature enabling model use in Guild AI. Without formally defining
these, users are left to their own to infer model capabilities from
available documentation and source code.

@@

A user may use the \verb|guild help| command to get details about a
model, its operations, operation flags, and resource
dependencies. This is similar to the \emph{man} facility (ref) on many
systems.

A user may also get help for a particular operation by using the
\verb|--help-op| option with the \verb|guild run| command. This is
similar to the \verb|--help| option available for many programs.

Figure~\ref{fig:op-help} contains sample help for the \verb|train|
operation of the \verb|gpkg.mnist/cnn| package model.

\begin{figure}
  \begin{lstlisting}[]
  Usage: guild run [OPTIONS] cnn:train
  [FLAG]...

  Train the CNN

  Use 'guild run --help' for a list of
  options.

  Dependencies:
    mnist-dataset  Yann Lecun's MNIST
                   dataset in compressed
                   IDX format

  Flags:
    batch-size  Number of images to
                include in a training
                batch (100)
    epochs      Number of epochs to train
                (10)
  \end{lstlisting}
  \caption{Sample operation help}
  \label{fig:op-help}
\end{figure}

It is important to note that model and operation help is automatically
generated from model descriptions. This facility lowers the cost of
model documentation for developers and gives users a consistent
interface for discovering model capabilities and intructions for using
them.

@@

Once a model is available---either as an installed package model or
defined in a project---it can be used by running \emph{operations}. An
operation is an action that can be performed on a model. Operations
both \emph{define and implement} model capabilities. The act of using
a model is the process of running operation.

Operations may define an entire application development life cycle for
one model or a host of models that work together.

Section~\ref{sec:operations} lists common operation names.

The act of running an operation starts a \emph{run}, which is both an
operating process and a Guild AI managed file system artifact that
contains run details including generated output.

A user can run an operation using the \verb|guild run| command.

Refer to Section~\ref{sec:runs} for information about managing runs.

@@

As users and developers run model operations, they can view the
results in various ways. Guild AI provides a comprehensive command
line interface for managing runs, which is outlined in
Section~\ref{sec:runs}.

Guild AI also provides various methods for visualizing runs. Guild
View is an application that is provided with Guild AI that launches a
web application that can be used to view run details, including run
files.

Guild AI also integrates TensorBoard (ref) as a visualization tool,
which can be used to view TensorFlow event logs that are generated by
model operations.\footnote{Operations must explicitly write TensorFlow
  event logs during a run to support TensorBoard visualization.}

\subsection{Accessing Generated Artifacts}

Operations generate \emph{artifacts}, which are Guild AI managed files
stored on the local file system. Guild AI does not attempt to
obfuscate generated artifacts---they are maintained in the run
directory as written during the operation.

Artifacts are routinely used as inputs to subsequent operations. For
example, a TensorFlow checkpoint generated by a \verb|train| operation
may be used as input to an \verb|export-and-freeze| operation, which
in turn generates a frozen inference graph. This process is automatic
and part of Guild AI's resource resolution process.

Artifacts may also be accessed directly by users given their file
system paths. A user can list the full paths for all files associated
with a run using the \verb|guild runs info| command with the
\verb|-ff| option. With these path, the user may copy the files or
link to them as needed.

Runs may also be \emph{exported} to a location using the
\verb|guild export| command. Exported runs are copied (or moved, if
the \verb|--move| option is specified) to a specified file system
directory. This can be used to create backups or to transfer runs to
another location.

@@

Information about one or more runs may be displayed using the
\verb|guild runs info| command. By default, high level run information
such as the run ID, operation, start time, command, exit status, and
operating system pid are displayed. Additional command options may be
used to view more run information including files, flag values,
dependencies, and operation output.

@@

Guild AI makes use of the Click Python library (ref) to implement its
core command line processing support.

@@

Guild AI packages are valid Python wheel distributions. The underlying
libraries used to install Guild packages are the same as those used by
the \verb|pip| program to install PyPI packages. As a result, Guild
AI can be used to install PyPI packages and pip can be used to install
Guild AI packages---the process is the same in both cases.

@@

Guild View is implemented using the VueJS library (ref) for the front
end and Werkzeug (ref) for the HTTP API back end.

@@

Guild AI is freely available under the Apache 2.0 open source
license. It has been under active development for the last two years
and undergone significant revision and updates during that time based
on user feedback and real world application development.

@@
